---
title: "Re-weighting methods and SDs in Continuous Outcomes"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    theme: readable
    css: "style.css"
    social: menu
    source: embed
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(flexdashboard)
library(svglite)
library(dplyr)
devtools::load_all()
```


### Out of balance
<font size = 20> <center> Single arm trials are often imbalanced on key characteristics </center> </font>
```{r ft.width = 8}
dat <- data.frame(trial = c(rep("A",3), rep("B",3)),
                  char = rep(c("sex","recep","age"),2),
                  val = c(30, 20, 40,
                          40, 40, 50),
                  sd = c(NA, NA, 8, NA, NA, 5))


format <- dat %>%
  dplyr::mutate(
      val = dplyr::case_when(char == "age" ~ glue::glue("{val} ({sd})"),
                           TRUE ~  glue::glue("{val}%"))
  ) %>% dplyr::select(-sd) %>%
  tidyr::pivot_wider(names_from = "trial", values_from = "val") %>%
  dplyr::mutate(
    char = dplyr::case_when(char == "sex" ~ "Sex (Male)",
                            char == "recep" ~ "Receptor Positive",
                            TRUE ~ "Age (Mean, SD)"))

format %>%
  flextable::flextable() %>%
  flextable::set_header_labels(
   char = "Characteristics",
   A = "Trial A",
   B = "Trial B") %>%
  flextable::fontsize(size = 25, part = "all") %>%
  flextable::set_table_properties(layout = "autofit") %>%
  flextable::theme_zebra(
    odd_body = "#fff5e4",
    odd_head = "#f1e8e3"
  ) %>%
  flextable::border_outer()

```

------

Single arm trials are increasingly common in oncology and other therapeutic areas.
Since these trials don't have a randomized comparator, treatment comparisons between
two or more therapies usually need to adjust for differences in patient characteristics
that can predict outcome. Propensity score methods, especially MAIC, are a common
approach and their use is straightforward when characteristics just require 
balance on percentages (eg, percent males). When patient characteristics are continuous, we
need to consider whether to adjust to just the mean or the mean and standard
deviation. 

### What are the odds?


```{r, fig.width= 9, dev = "svglite"}

sds <- seq(from = 0, to = 10)
n <- 100000 # Sidestep some randomness
p <- purrr::map_dbl(sds, ~ {
x <- rnorm(n,0, .)

round(mean(rbinom(n,1,plogis(qlogis(0.2) + log(2)*x)) == 1) * 100, 2)
  
})

pdat <- data.frame(p,sds)

pdat %>%
  ggplot2::ggplot(ggplot2::aes(x = sds, y = p)) +
  ggplot2::geom_point(size = 4, alpha = 0.9, colour = "black",fill = "#5070A2", pch= 21) +
  ggplot2::theme_minimal(base_size = 16) +
  ggplot2::labs(
    y = "Percent with response",
    x = "Standard deviation of continuous characteristic",
    title = stringr::str_wrap("In logistic regression the percentage of patients with a binary outcome changes if the standard
    deviation of a continuous characteristic changes", 70)
  )
```

------------------------------------------------------------------------

For binary outcomes like objective response at a given timepoint we will usually
assume assume that patient characteristics change the odds of response. This is
mostly a mathematical convenience since it ensures predictions will always be
valid probabilities (relative risks and risk differences will sometimes predict
impossible values). A drawback of odds ratios (and hazard ratios) is that they
are non-linear transformations which in our case means that if we really think the effect
of x is linear on the logit scale (or log hazard) the mean probability of response
will depend on both the mean and standard deviation of continuous predictors. 

### ITC Implications?

```{r dev = "svglite", fig.width = 9}
p5 <- pdat %>% dplyr::filter(sds == 5) %>% dplyr::pull(p)
n <- 300
pdat %>%
  dplyr::mutate(p = p /100,
                p_comp = p5/100,
                expa = n*p,
                expb = n-expa,
                expc = n*p_comp,
                expd = n-expc,
                var = 1/expa + 1/expb + 1/expc + 1/expd,
                or = qlogis(p) - qlogis(p_comp),
                upr = or + 1.96 * sqrt(var),
                lwr = or - 1.96 * sqrt(var)) %>%
  dplyr::select(p, p_comp,or, upr, lwr,sds) %>%
  ggplot2::ggplot(ggplot2::aes(x = sds, y = or, ymin = lwr, ymax = upr)) +
  ggplot2::geom_hline(yintercept = 0, linetype = 2) +
  ggplot2::geom_pointrange(colour = "#375180") +
  ggplot2::coord_flip() +
  ggplot2::scale_y_continuous(labels = function(x) round(exp(x),2)) +
  ggplot2::theme_minimal(base_size = 16) +
  ggplot2::labs(x = stringr::str_wrap("Standard deviation of continuous characteristic in your trial",40),
                y = stringr::str_wrap("Odds Ratio and 95% CI  (Your Drug vs Comparator)"),
                caption = "Confidence intervals are illustrative and based of observed outcome with n = 300 for each group",
                title = stringr::str_wrap("Reweighting methods can come to different conclusions adjusting to mean only in nonlinear models", 60)) +
  ggplot2::annotate(geom= "label", x = 5.5, y = log(0.61),label = "Comparator SD = 5",
                    fill = "#375180",
                    colour = "white")
  
```

------------------------------------------------------------------------

If both treatments have the exact same treatment effect the failure to
consider adjustment of both means and standard deviations can lead to conclusions
that will be biased if the outcome model is nonlinear (eg, odds ratios, hazard ratios).
Remember that the goal of propensity based methods is balance, and so we should
consider [balance of higher-moments in addition to means](https://onlinelibrary.wiley.com/doi/10.1002/sim.6607)


